---
title: "Beating the Clock, A Predictive Approach to First Responder Response Time"
author: "Samantha Doyle & Alyssa Holmes"
format: html
code-overflow: wrap
editor: visual
---

## Data Preparation

### Data Extraction

This data was sourced from *NYC Open Data*.

```{r}
# install.packages("readxl")
library(readxl)

# Import data
responder <- read.csv('End_to_end_911.csv')
head(responder)
attach(responder)
```

### Data Cleaning

The original dataset contains one date variable, two categorical variables, and twenty-seven numeric variables. All the numeric variables are measured in seconds except for X..of.Incidents.Calculated which counts the number of incidents in a given month. The two categorical variables are agency, which states the department of the emergency service, and Final.Incident.Type, which states the final incident category. Twelve numeric variables were department specific (i.e. Fire Department) measurements. Since the goal of this model is to predict emergency services as a whole, these variables were removed. A few of the remaining numeric columns had a singular missing value, which was imputed with the mean. A unique ID variable was created for each observation, as well. From the date variable, the month was extracted and sorted into categories by season (three months for spring, summer, fall, and winter each) to obtain another categorical variable. Finally, the missing values in Final.Incident.Type were imputed with the mode, and the 23 categories in this variable were condensed into a new categorical variable called IncidentClassification. 

```{r}
# Remove unwanted variables
responder_clean <- subset(
  responder,
  select = -c(
    Call.to.FDNY.Job.Creation,
    Call.to.FDNY.Pickup,
    Call.to.EMS.Pickup,
    Call.to.Agency.Job.Creation,
    Median.FDNY.Pickup,
    Median.FDNY.Job.Creation,
    Median.Ems.Pickup,
    Median.Ems.Job.Creation,
    Average.EMS.Pickup,
    Average.EMS.Processing,
    Average.FD.Processing,
    Average.FD.Pickup
  )
)

head(responder_clean)
summary(responder_clean)

# Find columns with missing values
colSums(is.na(responder_clean))

# Replace Average.Calltaker.Processing NA with column mean
responder_clean$Average.Calltaker.Processing[is.na(responder_clean$Average.Calltaker.Processing)] <-
  mean(responder_clean$Average.Calltaker.Processing, na.rm = TRUE)

# Replace Average.Dispatch NA with column mean
responder_clean$Average.Dispatch[is.na(responder_clean$Average.Dispatch)] <-
  mean(responder_clean$Average.Dispatch, na.rm = TRUE)

# Replace Median.Dispatch NA with column mean
responder_clean$Median.Dispatch[is.na(responder_clean$Median.Dispatch)] <-
  mean(responder_clean$Median.Dispatch, na.rm = TRUE)

# Replace Median.Calltaker.Handoff NA with column mean
responder_clean$Median.Calltaker.Handoff[is.na(responder_clean$Median.Calltaker.Handoff)] <-
  mean(responder_clean$Median.Calltaker.Handoff, na.rm = TRUE)

# Replace Call.to.PD.Calltaker.Handoff NA with column mean
responder_clean$Call.to.PD.Calltaker.Handoff[is.na(responder_clean$Call.to.PD.Calltaker.Handoff)] <-
  mean(responder_clean$Call.to.PD.Calltaker.Handoff, na.rm = TRUE)

# Recheck missing values
colSums(is.na(responder_clean))

# Create the ID variable
responder_clean$ID <- 1:nrow(responder_clean)

# Reorder the columns to place 'ID' at the beginning
responder_clean <- 
  responder_clean[, c("ID", setdiff(names(responder_clean), "ID"))]

# View the first few rows to check the result
head(responder_clean)

# Convert Week.Start.Date to Date format 
responder_clean$Week.Start.Date <- 
  as.Date(responder_clean$Week.Start.Date, format = "%m/%d/%Y")

# Create a new column for Season
responder_clean$Season <- ifelse(
  format(responder_clean$Week.Start.Date, "%m") %in% c("12", "01", "02"),
  "Winter",
  ifelse(
    format(responder_clean$Week.Start.Date, "%m") %in% c("03", "04", "05"),
    "Spring",
    ifelse(
      format(responder_clean$Week.Start.Date, "%m") %in% c("06", "07", "08"),
      "Summer",
      "Fall"  
    )
  )
)


# Reorder columns
responder_clean <- responder_clean[, c("ID", "Week.Start.Date", "Season", 
                                       setdiff(names(responder_clean), 
                                               c("ID", "Week.Start.Date", 
                                                 "Season")))]
# View data
head(responder_clean)

# Check for empty strings or zero-length entries
colSums(responder_clean == "")

# Convert empty strings to NA
responder_clean[responder_clean == ""] <- NA

# install.packages("dyplr")
library(dplyr)

# Reduce factor levels for Final.Incident.Type
responder_clean <- responder_clean %>%
  mutate(
    IncidentClassification = case_when(
      Final.Incident.Type %in% c(
        "1. Critical",
        "1. Life Threatening Med Emergencies",
        "1. Structural Fires"
      ) ~ "Critical",
      Final.Incident.Type %in% c(
        "2. Non-Life Threatening Med Emergencies",
        "2. Serious",
        "2. Non-Structural Fires"
      ) ~ "Serious",
      Final.Incident.Type %in% c("3. Medical Emergencies", "3. Non-Critical") ~
        "Non-Critical",
      Final.Incident.Type %in% c("4. Non-Medical Emergencies") ~ "Other",
      Final.Incident.Type %in% c("Alarms", "Shot Spotter") ~ "Alert",
      Final.Incident.Type %in% c("Disorderly", "Disorderly Person/Group/Noise",
                                 "Dispute") ~ "Dispute",
      Final.Incident.Type %in% c(
        "Hazardous Materials/Suspicious Letters/Packages/Substances/Substances",
        "United States Postal Service - Bio Hazard Detection System"
      ) ~ "Hazard",
      Final.Incident.Type %in% c(
        "Investigate/Possible Crime",
        "Other Crimes (In Progress)",
        "Possible Crimes"
      ) ~ "Possible Crime",
      Final.Incident.Type %in% c("Past Crime") ~ "Past Crime",
      Final.Incident.Type %in% c("Vehicle Accident") ~ "Vehicle Accident",
      TRUE ~ "Other",
      # If Incident.Type is NA, assign "Other"
    )
  )
```

## Exploratory Data Analysis

```{r}
# Reproducibility
set.seed(1)

# Make Train and Test using 80/20 split
sample <- sample(c(TRUE, FALSE), nrow(responder_clean),
                 replace=TRUE, prob=c(0.8,0.2))
train  <- responder_clean[sample, ]
test   <- responder_clean[!sample, ]

# Duplicate train data for further pre-processing
explore<- train

# Create histogram for target variable
hist(Call.to.Agency.Arrival, 
     main = "Histogram of Call to Agency Arrival", 
     xlab = "Call to Agency Arrival", 
     col = "purple", 
     border = "black")

##################################################
# The target variable is skewed right, indicating#
# the potential for log transformation           #
##################################################

# Create a new variable that is the log of 'Call.to.Agency.Arrival'
explore$Log.Call.to.Agency.Arrival <- 
  log(explore$Call.to.Agency.Arrival)

# Check the result by viewing the first few rows
head(explore[, c("Call.to.Agency.Arrival", "Log.Call.to.Agency.Arrival")])

# Check for NA values in the log-transformed column
sum(is.na(explore$Log.Call.to.Agency.Arrival))

# Check for NA values in the entire data frame
sum(is.na(explore))

# Histogram for log transformed target variable
hist(explore$Log.Call.to.Agency.Arrival, 
     main = "Histogram of Log (Call to Agency Arrival)", 
     xlab = "Call to Agency Arrival", 
     col = "purple", 
     border = "black")

# correlation matrix with log target variable
cor_matrix_log <- cor(explore[,sapply(explore,is.numeric)])


# install.packages("corrplot")
library(corrplot)

# Save the plot with larger dimensions
png("corrplot_log.png", width = 4000, height = 3200)  
corrplot(cor_matrix_log, method = "color", type = "upper", 
         addCoef.col = "black",  
         number.cex = 4,      
         tl.cex = 4)           
# Close the plotting device
dev.off()  

# Define the dataset variables (excluding the first three)
variables <- c(
  "X..of.Incidents.Calculated",
  "Call.to.First.Pickup",
  "Call.to.PD.Calltaker.Handoff",
  "Call.to.Agency.Dispatch",
  "Call.to.Agency.Arrival",
  "Call.to.First.Arrival..Multi.Agency.Incidents.",
  "Median.Pickup",
  "Median.Calltaker.Handoff",
  "Median.Dispatch",
  "Median.Travel",
  "Median.Cumulative.First.Arrival...Multi.Agency.Incidents.",
  "Average.Travel",
  "Average.Dispatch",
  "Average.Calltaker.Processing",
  "Average.Pickup",
  "Log.Call.to.Agency.Arrival"
)

# Loop through the variables and plot histograms for each variable
for (var in variables) {
  hist(explore[[var]], 
       main = paste("Histogram of", var), 
       xlab = var, 
       col = "purple", 
       border = "black")
}

##############################################
# Most variables are skewed right, indicating#
# the potential for log transformation       #
##############################################

# Define the dataset variables
log_variables <- c(
  "X..of.Incidents.Calculated",
  "Call.to.First.Pickup",
  "Call.to.PD.Calltaker.Handoff",
  "Call.to.Agency.Dispatch",
  "Call.to.Agency.Arrival",
  "Call.to.First.Arrival..Multi.Agency.Incidents.",
  "Median.Pickup",
  "Median.Calltaker.Handoff",
  "Median.Dispatch",
  "Median.Travel",
  "Median.Cumulative.First.Arrival...Multi.Agency.Incidents.",
  "Average.Travel",
  "Average.Dispatch",
  "Average.Calltaker.Processing",
  "Average.Pickup"
)

# Loop through the variables, take the log, and plot histograms
# Adding 1 to avoid log(0)
for (var in log_variables) {
  log_var <- log(explore[[var]] + 1)  
  
  # Create the histogram for the log-transformed variable
  hist(log_var, 
       main = paste("Histogram of Log", var), 
       xlab = paste("Log of", var), 
       col = "purple", 
       border = "black")
}

# Side-by-side boxplot of the log-transformed variable by Season
boxplot(Log.Call.to.Agency.Arrival ~ Season, 
        data = explore, 
        main = "Side-by-Side Boxplot by Season", 
        xlab = "Season", 
        ylab = "Log of Call to Agency Arrival",  
        col = "purple", 
        border = "black")

# Side-by-side boxplot of the log-transformed variable by Agency
boxplot(Log.Call.to.Agency.Arrival ~ Agency, 
        data = explore, 
        main = "Side-by-Side Boxplot by Agency", 
        xlab = "Agency", 
        ylab = "Log of Call to Agency Arrival",  
        col = "purple", 
        border = "black")

# Side-by-side boxplot of the log-transformed variable by Incident
boxplot(Log.Call.to.Agency.Arrival ~ IncidentClassification, 
        data = explore, 
        main = "Side-by-Side Boxplot by Incident Classification", 
        xlab = "Incident Classification", 
        ylab = "Log of Call to Agency Arrival", 
        col = "purple", 
        border = "black")

# ANOVA for Incident Classification
IC_Anova <- aov(Call.to.Agency.Arrival~IncidentClassification, data=train)
summary(IC_Anova)

IC_ttest<- TukeyHSD(IC_Anova)
IC_ttest

# ANOVA for Agency
Agency_Anova <- aov(Call.to.Agency.Arrival~Agency, data=train)
summary(Agency_Anova)

Agency_ttest<- TukeyHSD(Agency_Anova)
Agency_ttest

# List of variables to be log-transformed 
log_variables <- c(
  "X..of.Incidents.Calculated",
  "Call.to.First.Pickup",
  "Call.to.PD.Calltaker.Handoff",
  "Call.to.Agency.Dispatch",
  "Call.to.First.Arrival..Multi.Agency.Incidents.",
  "Median.Pickup",
  "Median.Calltaker.Handoff",
  "Median.Dispatch",
  "Median.Travel",
  "Median.Cumulative.First.Arrival...Multi.Agency.Incidents.",
  "Average.Travel",
  "Average.Dispatch",
  "Average.Calltaker.Processing",
  "Average.Pickup"
)

# Subset the data to include only numeric variables
log_data <- explore[, log_variables]

# Replace zero or negative values with 0.0001
log_data[log_data <= 0] <- 0.0001

# Log-transform the selected variables, adding 1 to avoid log(0)
log_data_transformed <- log(log_data + 1)

# Add the already log-transformed 'Log.Call.to.Agency.Arrival'
log_data_transformed$Log.Call.to.Agency.Arrival <- 
  explore$Log.Call.to.Agency.Arrival

# Numeric columns only
log_data_numeric <- 
  log_data_transformed[, sapply(log_data_transformed, is.numeric)]

# Create the correlation matrix
cor_matrix_log <- cor(log_data_numeric, use = "complete.obs")

# Save the plot with larger dimensions
png("corrplot_large_log.png", width = 4000, height = 3200)

# Generate the correlation plot
corrplot(cor_matrix_log, method = "color", type = "upper", 
         addCoef.col = "black", 
         number.cex = 4,         
         tl.cex = 4)             

dev.off()


##############################################
# Variables seem to be influenced by extreme #
# outliers; attempting winsorization         #
##############################################

# install.packages("ggplot2")
# install.packages("reshape2")
library(ggplot2)
library(reshape2)

# List of variables to winsorize
variables <- c( "X..of.Incidents.Calculated",
                "Call.to.First.Pickup",
                "Call.to.PD.Calltaker.Handoff",
                "Log.Call.to.Agency.Arrival",
                "Call.to.Agency.Dispatch",
                "Call.to.First.Arrival..Multi.Agency.Incidents.",
                "Median.Pickup",
                "Median.Calltaker.Handoff",
                "Median.Dispatch",
                "Median.Travel",
                "Median.Cumulative.First.Arrival...Multi.Agency.Incidents.",
                "Average.Travel",
                "Average.Dispatch",
                "Average.Calltaker.Processing",
                "Average.Pickup")

# Plot boxplots for all variables
explore_long <- log_data_transformed[variables] 

# Reshape data to long format for ggplot
explore_long <- melt(explore_long)  

ggplot(explore_long, aes(x = variable, y = value)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Boxplots Before Winsorization", x = "Variables", y = "Values")

# Winsorization function
winsorize <- function(x, lower_percent = 0.01, upper_percent = 0.99) {
  lower_limit <- quantile(x, lower_percent)
  upper_limit <- quantile(x, upper_percent)
  x[x < lower_limit] <- lower_limit
  x[x > upper_limit] <- upper_limit
  return(x)
}

# Apply winsorization to all variables
log_data_transformed[variables] <- lapply(log_data_transformed[variables],
                                          winsorize)

# Reshape data again for plotting
explore_long_winsorized <- log_data_transformed[variables]
explore_long_winsorized <- melt(explore_long_winsorized)

# Plot boxplots after winsorization
ggplot(explore_long_winsorized, aes(x = variable, y = value)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Boxplots After Winsorization", x = "Variables", y = "Values")
```

## Model Creation

### MLR Model

First, a standard MLR was created. Assumptions like normality and homosceadacity were tested. Unfortunately, this model did not pass the assumptions.

```{r}
# MLR Model
responder_MLR <- lm(
  Log.Call.to.Agency.Arrival ~
    X..of.Incidents.Calculated + Call.to.First.Pickup +
    Call.to.PD.Calltaker.Handoff + Call.to.Agency.Dispatch +
    Call.to.First.Arrival..Multi.Agency.Incidents. + Median.Pickup +
    Median.Calltaker.Handoff + Median.Dispatch + Median.Travel +
    Median.Cumulative.First.Arrival...Multi.Agency.Incidents. +
    Average.Travel + Average.Dispatch + Average.Calltaker.Processing +
    Average.Pickup + as.factor(Agency) + as.factor(IncidentClassification),
  data = explore
)

summary(responder_MLR)

# Assumption checking
plot(responder_MLR, which=1)
plot(responder_MLR, which=2)

# install.packages("lmtest")
library(lmtest)

# Breusch-Pagan (BP) test for heteroscedasticity
bptest(responder_MLR)

# Kolmogorov-Smirnov (KS) Test for normality of residuals
ks.test(residuals(responder_MLR), "pnorm", 
        mean = mean(residuals(responder_MLR)), 
        sd = sd(residuals(responder_MLR)))

##############################################
# Model did not pass; running model with     #
# log transformation                         #
##############################################
```

### Log Transformed MLR

Next, an MLR model was created using the log transformed variables. Assumptions were also checked for this model, but still it didn't pass.

```{r}
# Add categorical variables to the log transformed data
log_data_all <- log_data_transformed %>%
  mutate(Agency = explore$Agency,
         IncidentClassification = explore$IncidentClassification)

# MLR model with log transformed variables
responder_MLR_log <- lm(
  Log.Call.to.Agency.Arrival ~
    X..of.Incidents.Calculated + Call.to.First.Pickup +
    Call.to.PD.Calltaker.Handoff + Call.to.Agency.Dispatch +
    Call.to.First.Arrival..Multi.Agency.Incidents. + Median.Pickup +
    Median.Calltaker.Handoff + Median.Dispatch + Median.Travel +
    Median.Cumulative.First.Arrival...Multi.Agency.Incidents. +
    Average.Travel + Average.Dispatch + Average.Calltaker.Processing +
    Average.Pickup + as.factor(Agency) + as.factor(IncidentClassification),
  data = log_data_all
)

summary(responder_MLR_log)

# Assumption checking
bptest(responder_MLR_log)

ks.test(residuals(responder_MLR_log), "pnorm", 
        mean = mean(residuals(responder_MLR_log)), 
        sd = sd(residuals(responder_MLR_log)))

##############################################
# Model did not pass; running model with     #
# BoxCox transformation                      #
##############################################
```

### BoxCox Transformed MLR

Since the models were not passing the KS test, a BoxCox transformation was attempted.

```{r}
library(MASS)

# Apply BoxCox Transformation
boxcox_results <- boxcox(responder_MLR_log, lambda = seq(-2, 2, by = 0.1))

# Identify the best lambda value
lambda_best <- boxcox_results$x[which.max(boxcox_results$y)]
lambda_best

# Apply transformation to the response variable only
if (lambda_best > 0) {
  responder_clean$Transformed.Call.to.Agency.Arrival <- 
    (responder_clean$Call.to.Agency.Arrival ^ lambda_best - 1) / lambda_best
} else if (lambda_best == 0) {
  responder_clean$Transformed.Call.to.Agency.Arrival <- 
    log(responder_clean$Call.to.Agency.Arrival)
} else {
  responder_clean$Transformed.Call.to.Agency.Arrival <- 
    -1 / (responder_clean$Call.to.Agency.Arrival ^ lambda_best)
}

# MLR model with BoxCox-transformed response
responder_MLR_boxcox <- lm(
  Transformed.Call.to.Agency.Arrival ~
    X..of.Incidents.Calculated + Call.to.First.Pickup +
    Call.to.PD.Calltaker.Handoff + Call.to.Agency.Dispatch +
    Call.to.First.Arrival..Multi.Agency.Incidents. + Median.Pickup +
    Median.Calltaker.Handoff + Median.Dispatch + Median.Travel +
    Median.Cumulative.First.Arrival...Multi.Agency.Incidents. +
    Average.Travel + Average.Dispatch + Average.Calltaker.Processing +
    Average.Pickup + as.factor(Agency) + as.factor(IncidentClassification),
  data = responder_clean
)

summary(responder_MLR_boxcox)

# Assumption checking
bptest(responder_MLR_boxcox)

ks.test(residuals(responder_MLR_boxcox), "pnorm", 
        mean = mean(residuals(responder_MLR_boxcox)), 
        sd = sd(residuals(responder_MLR_boxcox)))

##############################################
# Model did not pass; running model with     #
# Winsorizing transformation                 #
##############################################
```

### Winsorized Model

After viewing the residual vs fitted plots and seeing the potential for outliers, a model was created using the winsorized data set.

```{r}
# MLR Model after Winsorization
responder_MLR_winz <- lm(Log.Call.to.Agency.Arrival ~ 
                       X..of.Incidents.Calculated+
                       Call.to.First.Pickup+
                       Call.to.PD.Calltaker.Handoff+
                       Call.to.Agency.Dispatch+
                       Call.to.First.Arrival..Multi.Agency.Incidents.+
                       Median.Pickup+
                       Median.Calltaker.Handoff+
                       Median.Dispatch+
                       Median.Travel+
                       Median.Cumulative.First.Arrival...Multi.Agency.Incidents.+
                       Average.Travel+
                       Average.Dispatch+
                       Average.Calltaker.Processing+
                       Average.Pickup,
                     data = log_data_transformed[variables])
summary(responder_MLR_winz)

# Assumption checking
bptest(responder_MLR_winz)

ks.test(residuals(responder_MLR_winz), "pnorm", 
        mean = mean(residuals(responder_MLR_winz)), 
        sd = sd(residuals(responder_MLR_winz)))


##############################################
# Model did not pass after winsorizing data; #
# checking now for multicollinearity         #
##############################################
```

### Addressing Multicollinearity

At this point, the models continue to fail the assumption check so multicollinearity was examined.

```{r}
# Assess multicollinearity 
eigen(cor(log_data_numeric))$values
max(eigen(cor(log_data_numeric))$values)/min(eigen(cor(log_data_numeric))$values)
kappa(cor(log_data_numeric), exact = TRUE)

model <- lm(Log.Call.to.Agency.Arrival~., data = log_data_numeric)

# install.packages("car")
library(car)

# Assess variance inflation factor
vif(model) 
mean(vif(model))

# Remove variables with VIF greater than 55 and refit the model
vif_MLR <- lm(Log.Call.to.Agency.Arrival ~ 
                X..of.Incidents.Calculated + Call.to.First.Pickup + 
                Median.Pickup + Average.Pickup+ Median.Calltaker.Handoff+
                Median.Dispatch+ Call.to.PD.Calltaker.Handoff+Median.Travel+
                Median.Cumulative.First.Arrival...Multi.Agency.Incidents.,
              data = log_data_numeric)
summary(vif_MLR)

vif(vif_MLR)

# Assumption checking
bptest(vif_MLR)

ks.test(residuals(vif_MLR), "pnorm", 
        mean = mean(residuals(vif_MLR)), 
        sd = sd(residuals(vif_MLR)))

##############################################
# Model did not pass with log transformation;#
# attempting GLM with gamma distribution and #
# log link                                   #
##############################################
```

### Gamma Distributed GLM

After considering that the MLR is not able to work for the data, a new approach, a Generalized Linear Model with a Gamma distribution and log link, was taken.

```{r}
# Gamma general linear model
glm_gamma <- glm(
  Call.to.Agency.Arrival ~ X..of.Incidents.Calculated + 
    Call.to.First.Pickup + Call.to.PD.Calltaker.Handoff + 
    Call.to.Agency.Dispatch +
    Median.Pickup +
    Median.Calltaker.Handoff + Median.Dispatch +
    Average.Calltaker.Processing +
    as.factor(Agency),
  data = train,
  family = Gamma(link = "log")
)

summary(glm_gamma)

# Assumption checking
bptest(glm_gamma)

ks.test(residuals(glm_gamma), "pnorm", 
        mean = mean(residuals(glm_gamma)), 
        sd = sd(residuals(glm_gamma)))

# Residuals vs Fitted Values (Homoscedasticity & Linearity)
plot(glm_gamma, which = 1)

# Normal Q-Q Plot (Normality of residuals)
plot(glm_gamma, which = 2)
```

### Ridge & LASSO Regression Model

To compare the GLM model created with other models, Ridge and LASSO regression models were generated.

```{r}
# install.packages("glmnet")
library(glmnet)

# RIDGE
y <- train$Call.to.Agency.Arrival
train_matrix <- as.matrix(train[, c(
  "X..of.Incidents.Calculated",
  "Call.to.First.Pickup",
  "Call.to.PD.Calltaker.Handoff",
  "Call.to.Agency.Dispatch",
  "Median.Pickup",
  "Median.Calltaker.Handoff",
  "Median.Dispatch",
  "Average.Calltaker.Processing"
)])

# Dummy code variables & split
dummy_agency <- model.matrix(~ Agency - 1, data = responder_clean)
sample2 <- sample(c(TRUE, FALSE), nrow(dummy_agency),
                  replace=TRUE, prob=c(0.8,0.2))
dummy_train <- dummy_agency[sample, ]
dummy_test <- dummy_agency[!sample, ]

# Combine dummy and numeric variables
X <- cbind(train_matrix, dummy_train)

# Fit a Ridge regression model
ridge_model <- glmnet(X, y, alpha = 0, lambda = 0.1)
coef(ridge_model)

# Perform cross-validation to select lambda
cv_ridge <- cv.glmnet(X, y, alpha = 0)
plot(cv_ridge$glmnet.fit, xvar = "lambda", label = TRUE)

best_lambda_ridge <- cv_ridge$lambda.min
ridge_model_best <- glmnet(X, y, alpha = 0, lambda = best_lambda_ridge)
coef(ridge_model_best)

# Fit a LASSO regression model
lasso_model <- glmnet(X, y, alpha = 1, lambda = 0.1)
coef(lasso_model)

# Perform cross-validation to select lambda
cv_lasso <- cv.glmnet(X, y, alpha = 1)
plot(cv_lasso$glmnet.fit, xvar = "lambda", label = TRUE)

best_lambda_lasso <- cv_lasso$lambda.min
lasso_model_best <- glmnet(X, y, alpha = 1, lambda = best_lambda_lasso)
coef(lasso_model_best)
```

## Model Comparison

When considering these models, one way to consider which model is best is to compare the AIC and BIC values, and the RMSE values.

```{r}
# install.packages("Metrics")
# install.packages("caret")
library(Metrics)
library(caret) 

# Test sets
X_test2 <- as.matrix(test[, c(
  "X..of.Incidents.Calculated",
  "Call.to.First.Pickup",
  "Call.to.PD.Calltaker.Handoff",
  "Call.to.Agency.Dispatch",
  "Median.Pickup",
  "Median.Calltaker.Handoff",
  "Median.Dispatch",
  "Average.Calltaker.Processing"
)])

y_test <- test$Call.to.Agency.Arrival
dummy_inc_test <- cbind(X_test2, dummy_test)

# Predict with GLM on the test set
predictions <- predict(glm_gamma, newdata = test, type = "response")

# Make predictions with the best Ridge model
pred_ridge_test <- predict(ridge_model_best, newx = dummy_inc_test)

# Make predictions with the best Lasso model
pred_lasso_test <- predict(lasso_model_best, newx = dummy_inc_test)

# Calculate RMSE for all models
rmse_gamma <- rmse(test$Call.to.Agency.Arrival, predictions)
rmse_ridge <- rmse(y_test, pred_ridge_test)
rmse_lasso <- rmse(y_test, pred_lasso_test)

# Comparing model metrics
aic_gamma<- AIC(glm_gamma)
bic_gamma<- BIC(glm_gamma)

# Function to calculate AIC and BIC for glmnet models
calculate_AIC_BIC <- function(model, X, y) {
  # Predicted values from the model
  pred <- predict(model, newx = X)
  
  # Calculate residual sum of squares (RSS)
  rss <- sum((y - pred)^2)
  
  # Number of observations and number of coefficients
  n <- length(y)
  k <- sum(coef(model) != 0) # Number of non-zero coefficients
  
  # Calculate AIC and BIC
  aic <- n * log(rss / n) + 2 * k
  bic <- n * log(rss / n) + log(n) * k
  
  return(list(AIC = aic, BIC = bic))
}
# Calculate AIC and BIC for Ridge
ridge_metrics <- calculate_AIC_BIC(ridge_model_best, X, y)
print(ridge_metrics)

# Calculate AIC and BIC for LASSO
lasso_metrics <- calculate_AIC_BIC(lasso_model_best, X, y)
print(lasso_metrics)

# Create a data frame to hold the results
results_table <- data.frame(
  Model = c("GLM Gamma", "Ridge", "Lasso"),
  AIC = c(aic_gamma, ridge_metrics$AIC, lasso_metrics$AIC),
  BIC = c(bic_gamma, ridge_metrics$BIC, lasso_metrics$BIC),
  RMSE = c(rmse_gamma, rmse_ridge, rmse_lasso)
)
print(results_table)
```

Using the table above, it can be said that for this particular data set, the LASSO regression model is the best option. A visual representation of this can be seen from the graphs below as well.

```{r}
# Plot Predicted vs Actual values for GLM
plot(test$Call.to.Agency.Arrival, predictions,
     main = "GLM: Predicted vs Actual",
     xlab = "Actual",
     ylab = "Predicted",
     col = "purple",
     pch = 16)
# Add identity line
abline(0, 1, col = "black", lwd=3)

# Plot Predicted vs Actual values for Ridge
plot(
  y_test,
  pred_ridge_test,
  main = "Ridge: Predicted vs Actual",
  xlab = "Actual",
  ylab = "Predicted",
  col = "purple",
  pch=16)
# Add identity line
abline(0, 1, col = "black", lwd=3) 

# Plot Predicted vs Actual values for LASSO
plot(
  y_test,
  pred_lasso_test,
  main = "LASSO: Predicted vs Actual",
  xlab = "Actual",
  ylab = "Predicted",
  col = "purple",
  pch=16)
# Add identity line
abline(0, 1, col = "black", lwd=3) 
```

## Conclusion

Using the AIC value as the main indicator, the final model choice would be the LASSO regression model. LASSO regression is able to minimize independent variables that are not important. Below is included a visual of the features included in the LASSO regression model in order of their importance from most to least as indicated by the bar graph. Feature importance was also created for the GLM model since it handles the variables differently.

```{r}
# Feature importance (via coefficients)
coefficients <- summary(glm_gamma)$coefficients[, 1]
cat("Feature Importance:")

# Create a data frame for the coefficients
importance_df <- data.frame(
  feature = names(coefficients),
  importance = abs(coefficients)
)
# Sort the importance in descending order
importance_df <- importance_df[order(importance_df$importance,
                                     decreasing = TRUE), ]
# Plot GLM feature importance
ggplot(importance_df, aes(x = reorder(feature, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "purple") +
  coord_flip() +
  labs(
    title = "GLM (Gamma) Feature Importance",
    x = "Feature",
    y = "Importance (Absolute Coefficient Value)"
  ) +
  theme_minimal()

# Feature Importance for LASSO regression
# Get coefficients
coef_lasso <- coef(lasso_model_best)

# Convert to tidy data frame
coef_df <- as.data.frame(as.matrix(coef_lasso))
coef_df$Feature <- rownames(coef_df)
colnames(coef_df)[1] <- "Coefficient"

# Remove intercept and zero coefficients
coef_df <- coef_df[coef_df$Feature != "(Intercept)" & 
                     coef_df$Coefficient != 0, ]

# Plot LASSO feature importance
ggplot(coef_df, 
       aes(x = reorder(Feature, abs(Coefficient)),
           y = abs(Coefficient))) +
  geom_bar(stat = "identity", fill = "purple") +
  coord_flip() +
  labs(title = "LASSO Feature Importance",
       x = "Feature",
       y = "Importance (Absolute Coefficient Value)") +
  theme_minimal()
```
